A Hitchhiker's Guide to Reproducible Research" # replace with something serious

### 2021-05-17 manuscript.Rmd
changes

---
title: "A Hitchhiker's Guide to Reproducible Research" # replace with something serious
author: Aaron Peikert and Andreas M. Brandmaier 
output:
  bookdown::pdf_document2:
    keep_tex: TRUE
repro:
  packages:
    - tidyverse
    - usethis
    - gert
    - aaronpeikert/repro@7bfaf98
    - here
    - rstudio/webshot2@f62e743
    - targets
    - renv
    - slider
    - patchwork
    - knitr
    - pander
    - lavaan
    - furrr
    - future.batchtools
  scripts:
    - R/simulation.R
    - R/simulation_funs.R
    - R/link.R
csl: apa7.csl
numbered_headings: true
bibliography: ["references.bib", "packages.bib"]
---

@@

```{r setup, include=FALSE}
library(repro)
automate_load_packages()
automate_load_scripts()
source(here::here("R", "link.R"))
```

```{r}
# define a ürint method for objects of the class data.frame
knit_print.data.frame = function(x, ...) {
  res = paste(c('', '', kable(x)), collapse = '\n')
  asis_output(res)
}
# register the method
registerS3method("knit_print", "data.frame", knit_print.data.frame)
```

@@

A cornerstone of scientific progress is the possibility to critically assess the validity of peer scientists’ scientific conclusions.
Scientific claims should not be credible because of their originators' authority but by the transparency and replicability of their supporting evidence [@aac4716].
Here, *replicability* refers to the degree to which original studies and their derivatives yield consistent results.
In other words, a successful replication study finds a consistent effect with an original study in a new sample of data.
Enabling the assessment of scientific conclusions requires a systematic and rigorous approach to research.
Researchers across various fields have identified a *crisis* of non-replicability of findings [e.g. @aac4716]. <!--citation needed, e.g. [@aac4716]-->

<!-- define replicability and then... 
 define reproducibility -->
Even worse, researchers might often not even be able to achieve reproducibility---defined as the ability to obtain identical results from the *same* data with the *same* computer code.
Over the past decades, computational statistical methods have radically changed the ability of researchers from all disciplines to process and analyze data with complex statistical models.
With these advances came challenges that contribute to broader concerns of irreproducibility in the published literature, among them the lack of exact and unambiguous documentation of computational methods and each computational step taken.
With computationally intensive data analyses, access to the steps taken to process data and generate findings is as important as access to the data themselves [@vanlissa2020worcs].
Only with access to both data and procedures can we guarantee full computational reproducibility, meaning that the creations emanating from a given data analysis (a statistical result or an entire manuscript describing the results) can be exactly reproduced at any time in the future, by any person, and on any computer.

@@

1.	Multiple inconsistent versions of code and data; for example, the dataset has changed over time because outliers were removed at a later stage, or the code has changed because a bug was removed at some point in time;
2.	Copy-and-paste errors; for example, results are manually copied from a computing environment into a text processor, causing inconsistencies between the reported result and the reproduced result;
3.	Ambiguous order of code execution; for example, with multiple data and code files, it may be unclear which scripts should be executed in what order;
4.	Broken dependencies; for example, a given analysis depends on a software package that might not be available on a different computer or no longer exists.

@@

This tutorial guides the reader through each step of an open research project with an emphasis on guaranteeing reproducibility from the very first step on.
This paper was designed to help researchers to practically apply the TOP guidelines [except for number 8. which does not apply to single research projects, @nosekPromotingOpenResearch2015; @vanlissa2020worcs].
We start by sketching an initial idea in a repository shared with colleagues that potentially collaborate only remotely.
We continue with the creation of a concrete data analysis plan, which ideally is already implemented in code, and a dynamic report document that can be preregistered.
Once the empirical data has been collected, the results can be calculated and appear directly in the dynamically generated document, where ideally only the discussion section needs to be added before the whole archive can be permanently stored as a fully reproducible digital artifact using a unique DOI in a persistent location.

@@

We have marked those steps that can be omitted as "optional" in the section headings.
Table \@ref(tab:overview) shows all steps required to to create and open and reproducible research project as they are presented in the remainder of the manuscript.

```{r overview, echo=FALSE}
overview <-
  read_csv(
    here::here("data", "overview.csv"),
    col_types = "cc",
    col_names = c("Phase", "Steps"),
    trim_ws = FALSE
  )
pander::pander(
  overview,
  split.tables = Inf,
  split.cells = Inf,
  missing = "",
  keep.line.breaks = TRUE,
  style = "grid",
  justify = 'left',
  caption = "\\label{tab:overview} Overview over all steps to create and open and reproducible research project."
)
```

@@

```{r, eval=FALSE}
check_github()
```

If these commands detect that a given software requirement is not installed or set up properly, you will get a message with detailed instructions (tailored to your operating system) on how to set up your computer.
Please make sure that all three commands pass before you continue with the tutorial.
This process is only required once per computer and does not need to be repeated per project.

# Planning Phase

<!-- researcher should begin early to enjoy most benefits -->
Reproducibility is invaluable for effective collaboration---including with your "past self".
While you might be tempted to consider reproducibility only when you about to finish a project, upfront investments pay off through your increased productivity.
Therefore, we suggest starting early, ideally from the moment you decide to pursue a project to incorporate tools to help you achieve reproducibility gradually.
The first step towards reproducibility is to bundle all required files in one folder.
This way, you never lose anything, and you can easily share the files without missing anything.
<!-- Hands-on: Introducing RProjects-->
RStudio's _R Projects_ facilitate this concept of a project folder.
To create one, click on:

@@

## Markdown

<!--Keep notes.-->
 A great tool to capture ideas or meeting notes is Markdown.
 Markdown is a lightweight text format that is very close to plain text with a minimal subset of reserved symbols for formatting instructions.
 This way, Markup does not need any specialized software for editing, it is highly fool-proof (unlike for example LaTeX) and can be ported to various document formats, such as HTML Websites, a Microsoft Word document, a typeset PDF file (for example, via LaTeX journal templates), or a Powerpoint presentation.
We suggest to use Markdown for everything, starting from simple sketches of your ideas to your scientific manuscripts and presentations---and even your CV.
To create a Markdown document, click on:

@@

Machiavellism is higher in male persons.

@@

And you could commit it in Git.
For convenience, we use the [`gert`-package](https://github.com/r-lib/gert) [@R-gert], which wraps most git commands into R functions.
Alternatively, we suggest to either work directly in the terminal (for example, RStudio offers a command shell with various options accessible in the Menu Tools → Terminal), or work using the Git pane in RStudio that offers buttons for basic operations.
<!--I am leaning towards doing everything in R, not sure though-->
<!--For Git we maybe should use RStudio's Git pane?-->

@@

Multigroup Confirmatory Factor Analysis with equal loadings and intercepts

@@

```r
# -- type this on the R console --
gert::git_commit_all("switch to CFA to exclude bias in measurement as possible confounder")
```

@@

```{r idea-change-screenshot, include=FALSE}
# unfortunately css selectors won't work for pdf so we'll use png instead
idea_change <- here::here("images", "idea-change.png")
if(!file.exists(idea_change)){
  webshot2::webshot("https://github.com/aaronpeikert/repro-tutorial/commit/76f5d4e78b3d3bf22e84e9137756bc17767dff3b?diff=split",
                  file = idea_change,
                  selector = "div#diff-2f8beaa39e5c98706d8abb3361a8383776f8df8a195a21c0e6eeb25e5aa48f45", # only select file change
                  zoom = 2)# higher resolution
}
```

@@

<!--I am troubled by how we should recommend to learn Git-->
Git can be challenging to learn, because it requires a rather complex mental model.
However, there is a subsets of commands which we present here and in the section on [GitHub] (commit, push, pull) which enable you to utilize Git for simple version control.
There are many tutorials on learning Git and @vuorreCuratingResearchAssets2018 provide a good starting point.
However Git is designed for collaboration, we therefore recommend finding a partner to attend an online lecture or workshop to learn the ins and outs of Git.

@@

```{r, include=FALSE}
source(here::here("R", "simulation_funs.R"))
```

@@

There are probably few things as frustrating as realizing that the data that you gathered cannot answer the research question you had in mind.
A step to prevent unpleasant surprises is to simulate data.
The machine learning community has begun to routinely simulate data to test models and their implementation [@braiekTestingMachineLearning2020].
Beyond checking the technical side of the model, simulation is also beneficial on a conceptual level [@axelrodAdvancingArtSimulation1997].
Simulated data lets you verify that the data you expect to gather and the analysis you plan fit together.
Beyond ensuring that the code runs without technical error and it lets you form more realistic expectations about the estimates' correctness and precision.
No matter which analytic strategy you plan, it comes with assumptions and a simulation forces you to address these assumptions more directly than simply thinking about them.
If you identify unrealistic assumptions, you can either violate these assumptions and see if the analytic strategy is robust or modify the analytical approach to address them.
We will show an example using structural equation modeling (SEM) for our Machiavellism example.
However, the general process applies to any statistical model, but the simulation here is specific to our hypothetical research question.
One has to make many educated guesses to simulate data that closely matches your expectation.
These guesses strongly depend on the study you are planning, and there cannot be a one fits all solution.
The analytic strategy (i.e. SEM) often gives you a hint about how to simulate data.

If you are unfamiliar with the specifics of SEM, you may briefly skim the next sections and continue in section ["Randome Number Generation"](#prng).
To simulate data in SEM, one first specifies a model (typically in the form of a structural model and a measurement model), selects plausible population values for all parameters, computes the model-implied mean vector and covariance matrix, and then samples as many observations as needed from a multivariate normal distribution following with the given mean vector and covariance matrix.

@@

```{r loadings, echo=FALSE}
knitr::kable(t(data.frame(Item = as.character(1:9),
                        Loading = loadings_jones_paulhus)),
             caption = "Factor loadings from an Exploratory Structrural Equation Model as reported in Jones \\& Paulhus (2013)")
```

@@

```{r}
model_truth <-
"MACH =~ 0.27*x1 + 0.22*x2 + 0.28*x3 + 0.36*x4 + 0.41*x5 + 0.5*x6 + 0.43*x7 +
0.32*x8 + 0.36*x9
MACH ~ c(0, 0.2)*1
MACH ~~ c(1, 1)*MACH"
```

@@

Data simulation is particularly easy in lavaan syntax using the function `simulateData`.
More complicated simulations (for example, including missing data simulation) are possible with other packages, e.g., simsem [@simsem], or simstudy [@goldfeldSimstudyIlluminatingResearch2020].
The following code simulates data from the population model `model_truth` with 1000 cases in each group.

```{r, include=FALSE}
set.seed(1234)
```

```{r}
simulated <- lavaan::simulateData(
  model_truth,
  sample.nobs = c(1000, 1000),
  meanstructure = TRUE,
  standardized = TRUE
)
```

@@

```{r}
model <- "MACH =~ x1 + x2 + x3 + x4 + x5 + x6 + x7 + x8 + x9
MACH ~~ c(1, NA)*MACH
MACH ~ c(0, NA)*1 + c(zero, diff)*1"
fit <-
  lavaan::cfa(
    model,
    data = simulated,
    group = "group",
    meanstructure = TRUE,
    std.lv = TRUE
  ) 
```

Now, `lavaan` warns us that this model may not be identified.
We overlooked a crucial assumption of this model, that is, measurement invariance.
This assumption is also baked into our simulation of the data.
There is a good chance that this assumption is violated.
We, therefore, adapt our analysis to plan for this contingency.
After fitting the model with strong measurement invariance, we automatically free equality constraints, first on loadings, and then on intercepts based on the Lagrange Multiplier test.
The model is iteratively refit and improved until at most three loadings, or three intercepts are freed, or the Lagrange Multiplier test is not significant for the constraints on loadings or intercepts.
While reconsidering our assumptions, it is also unlikely that the data is normally distributed.
Not only would the actual items be Likert-like and therefore categorical, but also are the in @Jones2013 reported means not in the middle of the scale, thus suggesting some skewness to the items.
For simplicity, we do not simulate these likely data problems (but doing so would undoubtedly make the simulation more realistic).
Anyhow, remember that what we consider realistic for our case could be unfit for your case.

If you are satisfied with your assumptions for the data simulation, it helps to write a function that expresses them, like this one:
<!--Hands-on: build simple functions tailored for your analysis-->

```{r generate_data}
```

And to build another function that captures your planned analysis.
Note that this function is again composed of several functions, which you may inspect on `r link("https://github.com/aaronpeikert/repro-tutorial/blob/main/R/simulation_funs.R", "GitHub")`.

```{r planed_analysis}
```

```{r extract_results}
```

With its help we can simulate 1000 cases and fit our analysis.

```{r}
generate_data(1000) %>% 
  planed_analysis() %>% 
  extract_results()
```

Note, that this certainly results in different numbers each time we run it (see ["Randome Number Generation"](#prng)).
Of course it is already helpful to simulate a single dataset to test if the analysis works in principle, but repeating the analysis several times and vary parameters lets you verify more properties of your analysis.
First, you can check if the planned analysis recovers the simulated values (`std_estimate` should be close to the simulated mean difference of 0.2).
Second, it enables you to plan an appropriate sample size because you are able to quantify the sampling variability.

To complete a such a  Monte-Carlo simulation study we need a function that repeats the analysis for different conditions (here only for different n):

```{r simulation_study}
```

Such Monte-Carlo simulation can help you decide between several tradeoffs.
One historically important tradeoff is between sample size and power (the probability to detect an effect of given size).
More recently there has been growing interest in other resource relevant parameters as well [!citation], such as number of measurement occasions, reliability of measures etc.
To show that there can also be alternative approaches for sample size planning beyond null hypothesis testing, we choose to visualize the density distribution of the parameter of interest (latent mean difference by gender of Machiavellism) instead of power.
Such process still lets you formulate expectations for the precision of the analysis, e.g. with a sample size of 1000 per group 80% of the simulated standardized effect size fall into a range of +/- .1, but without necessarily rely on null hypothesis testing.
For simplicity we do not show the code for the simulation/graphic, but it is available in the accompanying [GitHub repository](https://github.com/aaronpeikert/repro-tutorial)[^gh-repo]:.

[^gh-repo]: https://github.com/aaronpeikert/repro-tutorial

```{r, include=FALSE}
res <- read_csv(here::here("data", "simulation_results.csv"), col_types = "ddddd")
```

```{r hash}
```

```{r, include=FALSE}
res_interval <- res %>%
  select(n_obs, std_estimate) %>% 
  filter(!is.na(std_estimate)) %>% 
  group_by(n_obs) %>% 
  summarise(interval = list(interval(std_estimate, seq(0.05, .2, 0.01))), 
            .groups = "drop", median = median(std_estimate)) %>% 
  unnest(c(interval)) %>% 
  mutate(width = upper - lower, perc = 1 - alpha) %>% 
  group_by(alpha) %>% 
  arrange(n_obs) %>% 
  mutate(across(c(width, upper, lower),
                ~slide_dbl(.x, mean, na.rm = TRUE, .before = 2, .after = 0, .complete = FALSE))) %>% 
  ungroup() %>% 
  arrange(perc)
```

```{r, include=FALSE}
choosen_perc <- .8
choosen_width <- .2
true_diff <- .2
est_diff <- res %>% filter(n_obs == max(n_obs)) %>% pull(std_estimate) %>% mean()
minn <- filter(res_interval, perc == choosen_perc, width < choosen_width) %>% 
  filter(n_obs == min(n_obs))
sim_n <- pull(res, n_obs)
sim_nmax <- max(sim_n)
sim_n_unique <- unique(sim_n)
sim_nstep <- sim_n_unique[2] - sim_n_unique[1]
```

```{r, include=FALSE}
plot_interval <- res_interval %>%
  ggplot(aes(
    n_obs,
    ymin = lower,
    ymax = upper,
    fill = perc,
    group = alpha
  )) +
  geom_ribbon() +
  geom_line(aes(y = median), color = "white") +
  geom_hline(yintercept = true_diff, linetype = "dotted") +
  scale_fill_viridis_c(
    begin = 1,
    end = 0,
    breaks  = 1 - seq(0.05, .2, 0.05),
    labels = . %>% scales::percent(accuracy = 1)
  ) +
  scale_y_continuous(n.breaks = 10) +
  scale_x_continuous(breaks = seq(0, sim_nmax, sim_nstep * 10)) +
  theme_minimal() +
  labs(x = "N per Group", y = "Standardized Effect Size") +
  theme(legend.title = element_blank()) +
  NULL
```

```{r, include=FALSE}
plot_width <- res_interval %>%
  ggplot(aes(n_obs, width, color = perc, group = perc)) +
  geom_line() +
  geom_point(
    data = minn,
    size = 3,
    color = "black",
    shape = 3
  ) +
  scale_color_viridis_c(
    begin = 1,
    end = 0,
    breaks = 1 - seq(0.05, .2, 0.05),
    labels = . %>% scales::percent(accuracy = 1)
  ) +
  theme_minimal() +
  scale_y_log10(breaks = seq(0, 0.8, .1)) +
  scale_x_continuous(breaks = seq(0, sim_nmax, sim_nstep * 10)) +
  labs(x = "N per Group", y = "Width of quantile\n(Standardized Effect Size)") +
  theme(legend.title = element_blank()) +
  NULL
```

```{r, include=FALSE}
simulation_plot <- plot_interval +
  theme(axis.text.x = element_blank(), axis.title.x = element_blank()) + 
  plot_width +
  plot_layout(nrow = 2, guides = 'collect') +
  plot_annotation(tag_levels = 'A')
```

```{r, simulation-plot, echo=FALSE, fig.cap="*Plot A* depicts the range of effect sizes we observed in the simulation, where the colour encodes the density. The light yellow interval shows where 80% of the estimates land. The dotted line represents the simulated true effect size; the white line shows the median observed effect sizes. *Plot B* represents the width of the interval shown in Plot A on a logarithmic scale. The cross shows the minimum required sample size so that 80% of the estimates land in a range of +/- .1.  "}
simulation_plot
```

@@

Strict reproducibility is at stake with any code that is based on the computer's pseudo random number generator (PRNG) because random numbers will almost always differ with each execution of the code.
To consistently get the same results while preserving the pseudo-random property of the generated numbers, it is good practice to set a random seed that sets the PRNG into a restorable state, in R code: `set.seed(1235)`.
It is a little more complicated to achieve uncorrelated random numbers in parallelized code (e.g., when the same code is run with different portions of data in a computer cluster).
The standard random number generator (RNG) produces high quality random numbers, but is not safe to use in parallelized code (e.g., when the same code is run with different portions of data in a computer cluster).
The problem is that concurrent random number streams are correlated, because they repeat after some time.
Therefore, if you set a seed for one worker and another seed for another worker, you could end up with the same random numbers but with a phase shift.
@lecuyerGoodParametersImplementations1999 proposed a more computationally expensive RNG, designed to generate many uncorrelated streams.
While this RNG is available as a low level interface in R (with `RNGkind("L'Ecuyer-CMRG")`), we rely on the excellent parallelization framework around the [`future`-package](futureverse.org)[^futureverse] to handle this for us across all workers:

@@

Currently, there are various templates available that guide researchers in developing data analysis plans for preregistration. 
Still, with any natural-language representation of an analysis plan, there remains ambiguity when turning the natural language description of a given data analysis into code.
Here, we discuss preregistration as code (PAC) with the R package repro, which leaves no room for ambiguity.
PAC essentially reverses the process of empirical research.
Before researchers gather data, they write a reproducible, dynamically generated manuscript including the sections Introduction, Methods, and Results---initially based on simulated data.
Then, the source code of this manuscript serves as an unambiguous description of the analysis plan and is published as a preregistration.

PAC works well within this reproducible framework.

First, the preregistration becomes a part of the reproducible repository.
It is part of the set of version-controlled files and provides transparency about when the preregistration file, data files, and code files were changed over the course of a research project.
If code was version-controlled and tagged (see Section [Git]) before data collection, all deviations from the preregistration or any post-hoc additions that were performed after data collection are transparent and are visible and comparable in the version history.

One way to ensure that you can expect your analysis to work with real data is to simulate data.
We, for example, simulated data for the Machiavellism subscale of the short dark triad (SD3) and analyzed this with structural equation modelling (SEM).
The simulation showed that we have to gather at least `r minn$n_obs` observations per group to archive enough precision and that our solution to a data problem we expect (insufficient measurement invariance) introduces a bias of `r round((est_diff - true_diff)*100/true_diff, 2)`%.
These results are also visualized in Figure \@ref(fig:simulation-plot).
If you went to the trouble to simulate data, we argue that you get a preregistration almost for free.
A preregistration in code describes sufficiently how you want to analyze the data, but ideally, you also convey the why.
Dynamic document generation combines code and text, enabling you to explain your reasoning.
See the section on [R Markdown] for an example of a dynamic document generation paradigm.
In fact, this manuscript and therefore the here presented Machiavellism example were preregistered as a full manuscript, and you can see the results on real data in section [Reproduce a Project].

@@

```bash
# this should be run in the terminal
make DOCKER=TRUE
```
The function `automate()` combines `automate_docker()` & `automate_make()`.

>>>>

---
title: "A Hitchhiker's Guide to Reproducible Research" # replace with something serious
author: Aaron Peikert and Andreas M. Brandmaier 
output: bookdown::pdf_document2
repro:
  packages:
    - usethis
    - gert
    - aaronpeikert/repro@7bfaf98
    - here
    - rstudio/webshot2@f62e743
    - targets
    - renv
  scripts:
    - R/lavaan_helper.R
    - R/simulation.R
    - R/link.R
csl: apa7.csl
numbered_headings: true
bibliography: ["references.bib", "packages.bib"]
---

@@

```{r setup, include=FALSE}
library(repro)
library(tidyverse)
automate_load_scripts()
source(here::here("R", "link.R"))
```

@@

A cornerstone of scientific progress is the possibility to critically assess the validity of peer scientists’ scientific conclusions.
Scientific claims should not be credible because of their originators' authority but by the transparency and replicability of their supporting evidence [@aac4716].
Here, *replicability* refers to the degree to which original studies and their derivatives yield consistent results.In other words, a succesful replication study finds a consistent effect with an original study in a new sample of data.
Enabling the assessment of scientific conclusions requires a systematic and rigorous approach to research.
Researchers across various fields have identified a *crisis* of non-replicability of findings [e.g. @aac4716]- <!--citation needed, e.g. [@aac4716]-->

<!-- define replicability and then... 
 define reproducibility -->
Even worse, researchers might often not even be able to achieve reproducibility---defined as the ability to obtain identical results from the *same* data with the *same* computer code.
Over the past decades, computational statistical methods have radically changed the ability of researchers from all disciplines to process and analyze data with complex statistical models.
With these advances came challenges that contribute to broader concerns of irreproducibility in the published literature, among them the lack of exact and unambiguous documentation of computational methods and each computational step taken.
With computationally intensive data analyses, access to the steps taken to process data and generate findings is as important as access to the data themselves.
Only with access to both data and procedures can we guarantee full computational reproducibility, meaning that the creations emanating from a given data analysis (a statistical result or an entire manuscript describing the results) can be exactly reproduced at any time in the future, by any person, and on any computer.
Only if both the data and code are shared, can we obtain a reproducible scientific product, facilitating assessment of its validity and later reuse.

@@

1.	Multiple inconsistent versions of code and data; for example, the dataset has changed over time because outliers were removed at a later stage, or the code has changed because a bug was removed at some point in time;
2.	Copy-and-paste errors; for example, results are manually copied from a computing environment into a text processor, causing inconsistencies between the reported result and the reproduced result;
.3.	Ambiguous order of code execution; for example, with multiple data and code files, it may be unclear which scripts should be executed in what order;
4.	Broken dependencies; for example, a given analysis depends on a software package that might not be available on a different computer or no longer exists.

@@

This tutorial guides the reader through each step of an open research project with an emphasis on guaranteeing reproducibility from the very first step on.
We start by sketching an initial idea in a repository shared with colleagues that potentially collaborate only remotely.
We continue with the creation of a concrete data analysis plan, which ideally is already implemented in code, and a dynamic report document that can be preregistered.
Once the empirical data has been collected, the results can be calculated and appear directly in the dynamically generated document, where ideally only the discussion section needs to be added before the whole archive can be permanently stored as a fully reproducible digital artifact using a unique DOI in a persistent location.

@@

@@

If any of these commands find that a given software requirement is not installed or set up properly, you will get a message with detailed instructions (tailored to your operating system) on how to set up your computer.
Please make sure that all three commands pass before you continue with the tutorial.
This process is only required once per computer and does not need to be repeated for each project.

# Organizing Files

<!-- researcher should begin early to enjoy most benefits -->
Reproducibility is invaluable for effective collaboration---including with your "past self".
While you might be tempted to consider reproducibility only when you about to finish a project, upfront investments pay off through your increased productivity.
Therefore, we suggest starting early, ideally from the moment you decide to pursue a project to incorporate tools to help you achieve reproducibility gradually.
The first step towards reproducibility is to bundle all required files in one folder.
This way, you never lose anything, and you can easily share the files without missing anything.
<!-- Hands-on: Introducing RProjects-->
RStudio's _R Projects_ facilitate this concept of a project folder. The `repro` package offers an R Studio template, which sets up a file structure that follows best practices and conventions.<!--AP this is from your thesis - can you please add citations as to which practices and conventions these are?-->
To create one, click on:

@@

## Dynamic Documents

<!--Keep notes.-->
 A great tool to capture ideas or meeting notes is Markdown.
 Markdown is a lightweight text format that is very close to plain text with a minimal subset of reserved symbols for formatting instructions.
 This way, Markup does not need any specialized software for editing, it is highly fool-proof (unlike for example LaTeX) and can be ported to various document formats, such as HTML Websites, a Microsoft Word document, a typeset PDF file (for example, via LaTeX journal templates), or a Powerpoint presentation.
 R Markdown is the smart combination of Markdown with dedicated code chunks that are evaluated to formatted text, statistical results, or images when the document is rendered to its final format
 We suggest to use Markdown for everything, starting from simple sketches of your ideas to your scientific manuscripts and presentations---and even your CV.
To create a Markdown document, click on:

@@

Machiavellianism is higher in male persons.

@@

And you could commit it in Git.
For convenience, we use the [`gert`-package](https://github.com/r-lib/gert) [@R-gert], which wraps most git commands into R functions.
Alternatively, we suggest to either work directly in the terminal (for example, RStudio offers a command shell with various options accessible in the Menu Tools -> Terminal), or work using the Git pane in RStudio that offers buttons for basic operations.
<!--I am leaning towards doing everything in R, not sure though-->
<!--For Git we maybe should use RStudio's Git pane?-->

@@

Multigroup Confirmatory Factor Analysis
First, establish strong measurement invariance using a CFI cutoff-criterion
Second, test for differences in latent factor means of Machiavellianism using a $\chi^2$ test

@@

```r
# -- type this on the R console --
gert::commit_all("switch to CFA to exclude bias in measurement as possible confounder")
```

@@

```{r idea-change-screenshot, include=FALSE}
# unfortunately css selectors won't work for pdf so we'll use png instead
idea_change <- here::here("images", "idea-change.png")
if(!file.exists(idea_change)){
  webshot2::webshot("https://github.com/aaronpeikert/repro-tutorial/commit/5b3f4641ff542158184f85d458880c35e8f09f3c?diff=split",
                  file = idea_change,
                  selector = "div#diff-2f8beaa39e5c98706d8abb3361a8383776f8df8a195a21c0e6eeb25e5aa48f45", # only select file change
                  zoom = 2)# higher resolution
}
```

@@

<!--I am troubled by how we should recommend to learn Git-->
Git can be challenging to learn, because it requires a rather complex mental model of maintaining potentially differing versions between different repositories (typically one central repository and one local repository per each author while each repository may have different branches).
However, there is a subsets of commands which we present here and in the section on [GitHub] (commit, push, pull) which enable you to use Git for simple version control.
There are many tutorials on learning Git and @vuorreCuratingResearchAssets2018 provide a good starting point.
However Git is designed for collaboration, we therefore recommend finding a partner to attend an online lecture or workshop to learn the ins and outs of Git.

@@

```{r, include=FALSE}
source(here::here("R", "simulation.R"))
```

@@

There are probably few things as frustrating as realizing that the data that you gathered cannot answer the research question you had in mind.
A step to prevent unpleasant surprises is to simulate data.
The machine learning community has begun to use simulated data to test models and their implementation [@braiekTestingMachineLearning2020].
Beyond checking the technical side of the model, simulation is also beneficial on a conceptual leven [@axelrodAdvancingArtSimulation1997].
Simulated data lets you verify that the data you expect to gather and the analysis you plan fit together.
No matter which analytic strategy you plan, it comes with assumptions and a simulation forces you to address these assumptions more directly than simply thinking about them.
If you identify unrealistic assumptions, you can either violate these assumptions and see if the analytic strategy is robust or modify the analytical approach to address them.
We will show an example using structural equation modeling (SEM) for our Machiavellism example.
However, the general process applies to any statistical model, but the simulation here is specific to our hypothetical research question.
One has to make many educated guesses to simulate data that closely matches your expectation.
These guesses strongly depend on the study you are planning, and there cannot be a one fits all solution.
The analytic strategy (i.e. SEM) often gives you a hint about how to simulate data.

If you are unfamiliar with the specifics of SEM, we you may skip the next sections and continue in section ["Randome Number Generation"](#prng).
To simulate data in SEM, one first specifies a model (typically in the form of a structural model and a measurement model), selects plausible population values for all parameters, computes the model-implied mean vector and covariance matrix, and then samples as many observations as needed from a multivariate normal distribution following with the given mean vector and covariance matrix.

@@

```{r loadings, echo=FALSE}
knitr::kable(data.frame(Item = 1:9,
                        Loading = loadings_jones_paulhus),
             caption = "Factor loadings from an Exploratory Structrural Equation Model as reported in Jones \\& Paulhus (2013)")
```

@@

@@

```{r, results='asis', echo=FALSE}
model_truth <- combine(
  measurement(
    "MACH",
    items("x", 9),
    loadings = round(loadings_jones_paulhus * 0.7, 2)
    ),
  intercepts("MACH", list(0, 0.2))
  )
cat("```r",
    model_truth,
    "```",
    sep = "\n")
model <- combine(measurement(
  "MACH",
  items("x", 9)),
  intercepts("MACH", list(0, "NA")))
```

Data simulation is particularly easy in lavaan syntax using the function `simulateData`.
More complicated simulations (for example, including missing data simulation) are possible with other packages, e.g., simsem [@simsem], or simstudy [@goldfeldSimstudyIlluminatingResearch2020].
The following code simulates data from the population model `model_truth` with 1.000 cases in each group.

```{r, results='hide'}
simulated <- lavaan::simulateData(
  model_truth,
  sample.nobs = c(1000, 1000),
  meanstructure = TRUE,
  standardized = TRUE
)
```

@@

```{r}
fit <-
  lavaan::cfa(
    model,
    data = simulated,
    group = "group",
    meanstructure = TRUE,
    std.lv = TRUE
  ) 
```

Now, `lavaan` warns us that this model may not be identified.
We overlooked a crucial assumption of this model, that is, measurement invariance.
This assumption is also baked into our simulation of the data.
There is a good chance that this assumption is violated.
We, therefore, adapt our analysis to plan for this contingency.
After fitting the model with strict measurement invariance, we automatically free equality constraints, first on loadings, and then on intercepts based on the Lagrange Multiplier test.
The model is iteratively refit and improved until at most three loadings, or three intercepts are freed, or the Lagrange Multiplier test is not significant for the constraints on loadings or intercepts.
While reconsidering our assumptions, it is also unlikely that the data is normally distributed.
Not only would the actual items be Likert-like and therefore categorical, but also are the in @Jones2013 reported means not in the middle of the scale, thus suggesting some skewness to the items.
For simplicity, we do not simulate these likely data problems (but doing so would undoubtedly make the simulation more realistic).
Anyhow, remember that what we consider realistic for our case could be unfit for your case.

If you are satisfied with your assumptions, it helps to write a function that expresses them, like this one:
<!--Hands-on: build simple functions tailored for your analysis-->

```{r}
generate_data
```

Note that this function is again composed of several functions, which you may inspect on `r link("https://github.com/aaronpeikert/repro-tutorial/blob/main/R/simulation.R", "GitHub")`.

With its help we can simulate 500 cases and fit our analysis.

```{r}
true_diff <- 0.2
model_truth <- combine(
  measurement(
    "MACH",
    items("x", 9),
    loadings = round(loadings_jones_paulhus * 0.7, 2)
  ),
  intercepts("MACH", list(0, true_diff)),
  variances("MACH", list(1, 1))
)

model_same <- 
  "MACH =~ x1 + x2 + x3 + x4 + x5 + x6 + x7 + x8 + x9
  MACH ~ c(0,0)*1
  MACH ~~ c(1, NA)*MACH"

model_differ <- 
  "MACH =~ x1 + x2 + x3 + x4 + x5 + x6 + x7 + x8 + x9
  MACH ~ c(0, NA)*1 + c(zero, diff)*1
  MACH ~~ c(1, NA)*MACH"
generate_data(500,
              model_truth,
              model_same,
              model_differ,
              list(std_estimate = get_std_estimate),
              group.equal = c("loadings", "intercepts"),
              mod_load = 3,
              mod_int = 3,
              std.lv = TRUE)
```

Note, that this certainly results in different numbers each time we run it.

```{r}
generate_data(500,
              model_truth,
              model_same,
              model_differ,
              list(std_estimate = get_std_estimate),
              group.equal = c("loadings", "intercepts"),
              mod_load = 3,
              mod_int = 3,
              std.lv = TRUE)
```

@@

Strict reproducibility is at stake with any code that is based on the computer's pseudo random number generator (PRNG) because random numbers will almost always differ with each execution of the code.
To consistently get the same results while preserving the pseudo-random property of the generated numbers, it is good practice to set a random seed that sets the PRNG into a restorable state, in R code: `set.seed(1235)`.
It is a little more complicated to achieve uncorrelated random numbers in parallelized code (e.g., when the same code is run with different portions of data in a computer cluster).
The standard random number generator (RNG) produces high quality random numbers, but has is not safe to use in parrallelized code (e.g., when the same code is run with different portions of data in a computer cluster).
The problem is that concurrent random number streams are correlated, because they repeat after some time.
Therefore, if you set a seed for one worker and another seed for another worker, you could end up with the same random numbers but with a phase shift.
@lecuyerGoodParametersImplementations1999 proposed a more computationally expensive RNG, designed to generate many uncorrelated streams.
While this RNG is available as a low level interface in R (with `RNGkind("L'Ecuyer-CMRG")`), we rely on the [`future`-package](futureverse.org)[^futureverse] to handles this for us across all workers:

@@

Currently, there are various templates available that guide researchers in developing data analysis plans for preregistration. 
Still, with any natural-language representation of an analysis plan, there remains ambiguity when turning the natural language description of a given data analysis into code.
Here, we discuss preregistration as code (PAC) with the R package repro, which leaves no room for ambiguity.
PAC essentially reverses the process of empirical research.
Before researchers gather data, they write a reproducible, dynamically generated manuscript including the sections Introduction, Methods, and Results --- initially based on simulated data.
Then, the source code of this manuscript serves as an unambiguous description of the analysis plan and is published as a preregistration.

PAC works well within this reproducible framework.
First, the preregistration becomes a part of the reproducible repository.
It is part of the set of version-controlled files and provides transparency about when the preregistration file, data files, and code files were changed over the course of a research project.
If code was version-controlled and tagged (see Section XXX) before data collection, all deviations from the preregistration or any post-hoc additions that were performed after data collection are transparent and are visible and comparable in the version history.

One way to ensure that you can expect your analysis to work with real data is to simulate data.
A data simulation allows you to verify that the code runs error-free and lets you form more realistic expectations about the estimates' correctness and precision.
We, for example, simulated data for the Machiavellism subscale of the short dark triad (SD3) and analyzed this with structural equation modelling (SEM).
The simulation showed that we have to gather at least `missing_rcode_simulation` observations per group to archive enough precision and that our solution to a data problem we expect (insufficient measurement invariance) introduces a bias of `missing_rcode_simulation`%.
If you went to the trouble to simulate data, we argue that you get a preregistration almost for free.
A preregistration in code describes sufficiently how you want to analyze the data, but ideally, you also convey the why.
Dynamic document generation combines code and text, enabling you to explain your reasoning.
See the section on [R Markdown] for an example of a dynamic document generation paradigm.

@@

```bash
# this should be run in the terminal
make DOCKER=TRUE
```
